---
title: 'Homework #6 Linear Models with Categorical Variables'
author: "Nathan Strozewski"
date: "2022-10-19"
output: html_document
---

# Part 1: Comparing Two Means

### Background:
    1. In a dataset from Scantlebury et al, the authors explored how caste and mass affected the energy level of naked mole rats.

```{r part_1z}

library(readr)
library(ggplot2)
library(dplyr)
library(broom)
library(ggpmisc)
library(emmeans)
library(car)

theme_set(theme_classic(base_size = 14))
```

## Part 1a: 

### Tasks: 
    1. Load the data and plotting it for initial inspection
    2. We are looking at caste as a predictor of mass

```{r part_1a}

moles <- read.csv("/Users/nathanstrozewski/Documents/Everything/Education/M.S. Biology UMB/Courses/003_F2022/Biological Data Analysis/Homework/Homework #6/Data/18e4MoleRatLayabouts.csv")

str(moles)
summary(moles)

ggplot(data = moles,
       mapping = aes(x = caste,
                     y = lnmass)) +
  geom_boxplot()
```
## Part 2b:

### Tasks:
    1. Fit and evaluate the model.
    2. Does it meet assumptions? Why or why not?
    
```{r part_2b}

library(performance)

mole_lm <- lm(lnmass ~ caste,
               data = moles)

check_model(mole_lm)
```

### Notes:
    1. Model-predicted lines resemble observed data somewhat. Not an exact match but I am not super-concerned.
    2. Data appears linear-ish; reference line is flat-ish (curved), but ultimately ends linear. Perhaps a bit non-linear?
    3. Similar interpretation for homogeneity of variance
    4. It is hard to see, but it seems that the data points all fall within the (very narrow) contour lines
    5. Residuals look normal
    
## Part 1c & d:

### Notes: 
    1. Compare the two castes log mass values.
    2. Are they different? 
    3. How would you discuss your conclusions?
    4. Make a plot that you would feel would be clear and compelling in a publication

```{r part_1c}
moles_with_log_lnmass <- moles %>% 
  mutate(log_lnmass = log(lnmass))

ggplot(data = moles_with_log_lnmass,
       mapping = aes(x = caste,
                     y = log_lnmass,
                     fill = caste)) +
  geom_violin() +
  geom_point(color = "black") +
  geom_jitter() +
  stat_summary(ffun.data = "mean_cl_boot",
               geom = "pointrange",
               colour = "black",
               width = 0.2) +
  labs(title = "Does Naked Mole Rat Caste Predict Mass?",
       subtitle = "A Logarithmic Comparison",
       caption = "Figure 1c: Mass (log) is greater in naked mole rats of the lazy caste compared to the worker caste",
       x = "Caste",
       y = "Mass (log)") +
  theme(legend.position = "none")
```

### Conclusion:
    1. The lazy caste has a larger log(lnmass) than the worker caste

# Part 2:

## Background:
   1. To start with, let’s warm up with a simple one-way ANOVA model
   2. An example from Whitlock and Schluter (chapter 15 question 22) looks at the mass of lodgepole pinecones from different habitats

## Part 2a:

### Tasks:
    1. Load and plot the data. 
    2. Choose a plot that not only shows the raw data, but also the means and SE or CI of those means.
    3. Impress yourself by making it super fancy! Geom to your heart’s content.

```{r part_2a}   
cones <- read.csv("/Users/nathanstrozewski/Documents/Everything/Education/M.S. Biology UMB/Courses/003_F2022/Biological Data Analysis/Homework/Homework #6/Data/15q22LodgepolePineCones.csv")

str(cones)
summary(cones)

ggplot(data = cones,
       mapping = aes(x = habitat,
                     y = conemass,
                     fill = habitat)) +
  geom_violin() +
  geom_point(color = "black") +
  geom_jitter() +
  stat_summary(fun = mean,
               geom = "point",
               color = "black") +
  stat_summary(fun.data = mean_sdl,
               geom = "errorbar") +
labs(title = "Does Pine Cone Mass Vary by Habitat?",
       caption = "Figure 2a: Island.absent pine cones are larger in mass than Island.present or Mainland.present",
       x = "Habitat",
       y = "Pine Cone Mass") +
  theme(legend.position = "none")
```

## Part 2b:

### Tasks:
    1. Fit a model using least squares and evaluate all relevant assumptions. 
    2. List them out as you test them. 
    3. Can we use this model? If not, fix it. But if we can, no fix is needed!

```{r part_2b_model}
cones_lm <- lm(conemass ~ habitat,
               data = cones)
```

```{r part_2b_predictions}
check_predictions(cones_lm) %>% 
  plot()
```

### Notes: Observed data fits the predicted data very closely

```{r part_2b_linearity}
plot(cones_lm, which = 1)
```

### Notes: Similar residual structures

```{r part_2b_homogeneity}
plot(cones_lm, which = 1)
```

### Notes:
    1. There is an even-ish spread dots along the Y axis for both groups. This suggests that the variance between groups is relatively even

```{r part_2b_normality}
check_normality(cones_lm) %>% 
  plot(type = "qq")
```
### Notes:
    1. The dots are mostly near the fitted line!

```{r part_2b_outliers}
check_outliers(cones_lm) %>% 
  plot(type = "bar")
```
### Notes:
    1. No observations cross the threshold of 1

### Conclusions:
    1. I can continue with this model with no modifications

## Part 2c:

### Tasks:
    1. How much variation is explained by your model?
    
```{r part_2c}
summary(cones_lm)
```

### Conclusions:
    1. Multiple R-squared is 0.8851, while adjusted R-squared is 0.8675

## Part 2d: 

### Tasks:
    1. Show which means are different from each other.
    2. What CI are you using? 
    3. Are you adjusting your CIs? 
    4. Justify these choices.

```{r part_2d_emmeans}
cones_means <- emmeans(cones_lm, specs = ~habitat)
cones_means
```
### Notes:
    1. island.absent seems unique while island.present and mainland.present seem to overlap
    2. I can verify this using contrast()

```{r part_2d_contrast}
cones_means_contrast <- contrast(cones_means, method = "pairwise")
cones_means_contrast
```
### Notes:
    1. island.absent has a positive estimate when compared to island.present and mainland.present, suggesting that the means are unique
    2. island.present has a negative estimate when compared to mainland.present, suggesting that these means are not unique
    3. This is further supported by a high p-value of 0.9925
    
```{r part_2d_confint}
confint(cones_means_contrast, level = 0.85)
```
### Notes:
    1. Nothing surprising here
    2. I can plot this to further verify

```{r part_2d_plot}
plot(cones_means_contrast, level = 0.85) +
  geom_vline(xintercept = 0, lty = 2, color = "red")
```

### Notes:
    1. This visually confirms what I saw previously
    2. I have adjusted the CI with little change. See proof below:
    
```{r part_2d_CI_85}
confint(cones_means_contrast, level = 0.85)
```

```{r part_2d_CI_99}
confint(cones_means_contrast, level = 0.99)
```

```{r part_2d_CI_10}
confint(cones_means_contrast, level = 0.10)
```

# Part 3:

## Background:
   1. In a study from Rogers et al. (2020) the authors performed an experiment where they moved panels that had been colonized by invertebrates on a dock to a nearby rocky jetty where predators could access panels. 
   2. To separate out the effects of changes in abiotic environment versus predation, they performed a factorial experiment, either caging or not caging panels and placing them either on the side of a cinder block or hanging on a piece of PVC attached to the block where predators would have little access (but weren’t entirely stopped). 
   3. They then looked at change in total cover of invertebrates. 

## Part 3a:

### Task:
    1. Load the data
    2. Combine cages and position into a single variable treatment
    3. Plot the data. Note: we are interested in change in percent cover as influenced by the combined influence of caging and position
    4. Choose a plot that not only shows the raw data, but also the means and SE or CI of those means
    5. Impress yourself by making it as fancy as possible!

```{r part_3a}

foul <- read.csv("/Users/nathanstrozewski/Documents/Everything/Education/M.S. Biology UMB/Courses/003_F2022/Biological Data Analysis/Homework/Homework #6/Data/fouling_transplant_data.csv")

str(foul)

foul$Treatment.New <- paste(foul$Treatment, foul$Caged)
str(foul)

part_3a <- ggplot(
  data = foul,
  mapping = aes(x = Treatment.New,
                y = Change.in.Cover,
                fill = Treatment.New)
) +
  geom_violin() +
  geom_point(color = "black") +
  geom_jitter() +
  stat_summary(fun = mean,
               geom = "point",
               color = "black") +
  stat_summary(fun.data = mean_sdl,
               geom = "errorbar") +
  labs(title = "Does Treatment Influence Percent Cover?",
       x = "Treatment",
       y = "Percent Cover") +
  theme(legend.position = "none")
part_3a
```

## Part 3b:

### Tasks:
    1. Fit a model looking at how treatment influences cover change and evaluate all relevant assumptions. 
    2. Do you meet assumptions?

```{r part_3b}
foul_lm <- lm(Change.in.Cover ~ Treatment.New,
              data = foul)
check_model(foul_lm)
```
### Notes:
    1. The observed and predicted data aren't a super close match
    2. The data is nearly linear but not quite
    3. The comparisons have very different variances
    4. No big issues with normality
    5. No big issues with outliers

### Conclusions:
    1. The data does not meet assumptions

## Part 3c:

### Notes:
    1. If you answered yes to the above...you are wrong. It doesn’t! Percentage data is weird. Difference in percentages can be ever weirder!

### Tasks:
    1. Divide change by initial cover to express change as percent change relative to initial cover.
    2. Calculate difference in logit cover (so, logist(initial cover) - logit(final cover)). 
    3. Which one works (or at least works better) so that you can produce valid inference?
    
```{r part_3c}
foul_part_3c_1 <- foul %>% 
  mutate(PCR_to_IC = Change.in.Cover / Initial.Cover)
foul_part_3c_1

foul_part_3c_2 <- foul_part_3c_1 %>% 
  mutate(logit_cover = logit(Initial.Cover) - logit(Final.Cover))
foul_part_3c_2
```
    
### Notes:
    1. The two methods produced different values. 
       a. logit calculations are larger than dividing change in cover by initial cover
       b. Positive values in the former are negative in the latter, and vice versa
       c. I guess this makes sense when you consider logs
    
## Part 3d:

### What does this fit model tell you about how and which treatments differ in change?

```{r part_3d}
part_3d <- ggplot(
  data = foul_part_3c_2,
  mapping = aes(x = Treatment.New,
                y = PCR_to_IC,
                fill = Treatment.New)
) +
  geom_violin() +
  geom_point(color = "black") +
  geom_jitter() +
  stat_summary(fun = mean,
               geom = "point",
               color = "black") +
  stat_summary(fun.data = mean_sdl,
               geom = "errorbar") +
  labs(title = "Does Treatment Influence Percent Cover?",
       x = "Treatment",
       y = "Percent Cover") +
  theme(legend.position = "none")
part_3d
```

### Notes:   
    1. There is a lot of overlap in percent cover across treatment types
    2. Caged samples don't vary much in percent cover, and a general increase or decrease in percent cover isn't obvious 
    3. Open samples exhibit a higher spread of percent cover. Perhaps open samples are more susceptible to predation, leading to variance towards a decrease in percent cover. Conversely, perhaps open samples are not limited in space (like caged samples) and can grow more freely in the absence of predation, leading towards an increase in percent cover.
    4. Another factor in susceptibility to predation is position on block. Let's plot that to see if this influences percent cover and supports my hypothesis:
    
### Does position on block influence percent cover?

```{r part_3d_block}
ggplot(data = foul_part_3c_2,
       mapping = aes(x = Position.On.Block,
                     y = PCR_to_IC,
                     fill = Position.On.Block)) +
  geom_violin() +
  geom_point(color = "black") +
  geom_jitter() +
  stat_summary(fun = mean,
               geom = "point",
               color = "black") +
  stat_summary(fun.data = mean_sdl,
               geom = "errorbar") +
  labs(title = "Does Position of Sample Influence Percent Cover?",
       x = "Position on Block",
       y = "Percent Cover") +
  theme(legend.position = "none")
```

### Notes:
    1. Most samples are generally unaffected by position on block, though there are several samples demonstrating a substantial decrease in percent cover when placed on the side of the block (more susceptible to predation). These data are inconclusive but suggest there may be something here to dig into further.

# Part 4: Meta Questions

## Meta 1:

### Are you starting to see the connections between linear regression and models with categorical variables?
    1. Yes, it has clicked well for me
    2. These types of models plot really well for easy comparison, particularly with violin plots

### What about this week surprised you the most?
    1. Nothing particularly surprised me this week. I think the content flowed really well with that from last week. The two build well off of each other.

## Meta 2:

### What about what we did this week was the most difficult for you to understand? 
    1. I get the application of all the material and how to use it in R Studio to ask and answer questions. But I am still confused a bit by the statistical math that comprises the functions we use. The algebra/symbols are a bit foreign to me and will take time to get used to. It's like reading letters from a different alphabet

### Where are you feeling the boundaries of your understanding pushed the most?
    1. It is really easy for me to find one method of visualizing data that works ok-ish and stick with that. I need to continue pressing myself to formulate the best way to visualize each individual dataset to answer a question or demonstrate the findings

## Meta 3:

### How much time did this take you, roughly?
    1. 3.5 hours

## Meta 4:

### Please give yourself a weak/sufficient/strong assessment on this assigment. Feel free to comment on why.
    1. I always feel weird giving myself a strong assessment because I don't yet know if I actually did everything correctly.
    2. So I'll continue to give myself a strong-sufficient rating. I feel comfortable working through these question types, formulating approaches to address the questions, and implementing the solutions. Once I know whether or not my approach is correct, I'll shift my assessment to either strong or sufficient.



