---
title: "Tidyr Data Homework"
author: "Nathan Strozewski"
date: "2022-10-03"
output: html_document
---

# Part 1: "Load Me"

## Part 1a: Load libraries

```{r loading_libraries_1}

library(tidyr)
library(dplyr)
```

## Part 1b: Read data without downloading

```{r read_no_download}

library(readr)
kelp <- read_csv('https://github.com/kelpecosystems/observational_data/blob/master/cleaned_data/keen_cover.csv?raw=true')
```

## Part 1c: Read data with downloading

```{r read_with_download}

setwd("/Users/nathanstrozewski/Documents/Everything/Education/M.S. Biology UMB/Courses/003_F2022/Biological Data Analysis/Homework/Homework #3")
kelp <- read.csv("keen_cover.csv")
```

# Part 2: Format

## Part 2a: Is the data in the wide or long format?

```{r data_review}

str(kelp)
summary(kelp)
```
###
1. The data is in the long format
2. I can tell this because each row is an observation, while each column is a variable that data was collected for during each observation

# Part 3: Check it out

## Part 3a: How many sites has each PI done?

```{r PI_sites}

kelp_PIsites <-
  kelp %>% 
  group_by(PI) %>% 
  summarize(num_sites = n_distinct(SITE)) %>% 
  pivot_wider(names_from = PI,
              values_from = num_sites)
kelp_PIsites
```

## Part 3b: How many years of data does each site have? Show it in descending order.

```{r site_years}

kelp_yearsofdata <-
  kelp %>% 
  group_by(SITE) %>% 
  summarize(num_years = n_distinct(YEAR)) %>% 
  pivot_wider(names_from = SITE,
              values_from = num_years)
kelp_yearsofdata

kelp_yearsofdata_desc <- kelp_yearsofdata[order(kelp_yearsofdata, 
                                                decreasing = TRUE)]
kelp_yearsofdata_desc
```

## Part 3c: Impress yourself:
### Make a figure showing which site was sampled when 
### Use slice() (there are more elegant solutions)
### For data viz, you can use geoms you’ve used before, or new ones, like geom_tile()!

```{r site_sampling}

library(ggplot2)

kelp$SAMPLE_DATE <- as.Date(with(kelp, paste(YEAR, MONTH, DAY, sep = "-")), "%Y-%m-%d")
kelp$SAMPLE_DATE

kelp_sampledates <- ggplot(data = kelp) +
  labs(title = "Which sites were sampled -",
       subtitle = "and when?",
       x = "Sampling Site",
       y = "Date (YYYY-MM-DD") +
  geom_tile(aes(x = SITE,
                y = SAMPLE_DATE))
kelp_sampledates
```

# Part 4: Let's look at some kelp!

## Part 4a: Trim down to the cols YEAR, SITE, TRANSECT, PERCENT_COVER, FAMILY, and SPECIES

```{r kelp_select}

kelp %>% 
  select(YEAR, SITE, TRANSECT, PERCENT_COVER, FAMILY, SPECIES) 
```

## Part 4b: Trim the data down to only the family “Laminariaceae”. After that, ditch the FAMILY column

```{r kelp_Lami}

kelp %>% 
  filter(FAMILY == "Laminariaceae") %>% 
  select(YEAR, SITE, TRANSECT, PERCENT_COVER, SPECIES)
```

## Part 4c: For each species is there only one measurement per species transect each year? Or do we need to worry?

```{r kelp_sanity_check}

str(kelp) # stir the data to find out why
kelp[,13] # look at the size data to see if this explains why
```
### Answer:
    1. I conclude that the measurements are likely from both adult and juveniles

## Part 4d: Sum the cover for each species on each transect so that we only have one measurement per species (adults + juveniles together)

```{r kelp_transect_species_condense}

kelp_transectmean <- kelp %>% 
  # filter(!is.na(PERCENT_COVER)) %>% 
  group_by(SPECIES, TRANSECT) %>% # add site
  summarize(mean_transect_cover = mean(PERCENT_COVER, na.rm = TRUE)) %>% 
  print()

# unclear if I am supposed to have one measurement per transect per species ...
# or one measurement per species overall
# the former is one measurement per transect per species
# the following is one measurement per species

kelp_overallcover <- kelp %>% 
  group_by(SPECIES, SITE) %>% 
  summarize(mean_transect_cover = mean(PERCENT_COVER, na.rm = TRUE)) %>% 
  print()
```

## Part 4e: Make a plot showing the timeseries of kelps at each site. 
### You’ll want stat_summary() here. 
### You might even need it twice because ...
### note - stat_summary() has a geom argument where you can do things like “line”. 
### What might that do? Check it out! 
### Facet this plot by species, so we can see the trajectory of each. 
### Feel free to gussy this plot up however you would like (or not). 
### Do you notice anything? Comment!

```{r kelp_timeseries}

kelp_timeseries <- ggplot(data = kelp,
                          aes(x = SAMPLE_DATE,
                              y = PERCENT_COVER)) + # plot with values
  labs(title = "Kelp Cover Over Time",
       x = "Date (YYYY-MM-DD)",
       y = "Percent Cover") + # labels
  stat_summary(aes(x = SAMPLE_DATE,
               y = PERCENT_COVER), 
               geom = "line",
               width = 0.5,
               size = 0.5,
               alpha = 0.5) + # stats to summarize + dimensions
  scale_y_continuous(limits=c(0, 110)) # change y-axis limit
kelp_timeseries

kelp_timeseries +
  geom_jitter() + # adds jitter to allow plotting of overlapping data points
  facet_wrap(~ SPECIES) # splits plots into individ. plots per species
```
### Notes:
    1. I needed to add a jitter to plot the data, presumably because some points overlap
    2. It is quite difficult to make out a trend
    
# Part 5: Wide Relationships

## Part 5a: Using pivot_wider()
### Use pivot_wider() to make species into columns with percent cover as your values
### Note - be careful to fill in NAs as 0s

```{r kelp_pivotwider}
 
kelp_overallcover_wide <- kelp_overallcover %>% 
  pivot_wider(names_from = SPECIES,
                values_from = mean_transect_cover,
                values_fill = 0)
kelp_overallcover_wide
```

## Part 5b: Plot the relationship between Saccharina latissima and Laminaria digitata
### Add a line to your ggplot stat_smooth(method = "lm")
### Remember that you will need backticks ` around variables with spaces in them
### What do you think?
### Feel free to use any other geoms or explore however you like here

```{r kelp_sacc_lati_relationship}

ggplot(data = kelp_overallcover_wide,
         aes(x = SITE,
             y1 = get('Saccharina latissima'),
             y2 = get('Laminaria digitata'))) +
  geom_point(aes(x = SITE,
                 y = get('Saccharina latissima'),
                 color = "Saccharina latissima",
                 group = 1)) +
  geom_line(aes(x = SITE,
                y = get('Saccharina latissima'),
                color = "Saccharina latissima",
                group = 1)) +
  geom_point(aes(x = SITE,
                 y = get('Laminaria digitata'),
                 color = "Laminaria digitata",
                 group = 1)) +
  geom_line(aes(x = SITE,
                y = get('Laminaria digitata'),
                color = "Laminaria digitata",
                group =1)) +
  labs(title = "Mean Kelp Cover by Site",
       subtitle = "Saccharina latissima vs. Laminaria digitata",
       x = "Site",
       y = "Mean Percent Kelp Cover (%)",
       color = "Species")
```

## Part 5c: Pivot this correct long data back wide and then remake the figure from 4e
### Does it look different? Does it tell a different story?

### Notes:
    1. I am a little confused by the wording here
    2. The objects I am working with (the "correct" data) are in wide already
    3. Pivoting this long creates some issues
    4. Wouldn't it be easier to fill 0s for NAs in the original data and plot?

```{r kelp_timeseres_reevaluation}

# Replace NAs with base R

kelp_noNAs <- kelp
kelp_noNAs[is.na(kelp_noNAs)] = 0
kelp_noNAs

# Plot as in 4e

kelp_timeseries_noNAs <- ggplot(data = kelp_noNAs,
                          aes(x = SAMPLE_DATE,
                              y = PERCENT_COVER)) + # plot with values
  labs(title = "Kelp Cover Over Time",
       x = "Date (YYYY-MM-DD)",
       y = "Percent Cover") + # labels
  stat_summary(aes(x = SAMPLE_DATE,
                   y = PERCENT_COVER), 
               geom = "line",
               width = 0.5,
               size = 0.5,
               alpha = 0.5) + # stats to summarize + dimensions
  scale_y_continuous(limits=c(0, 110)) # change y-axis limit
kelp_timeseries_noNAs

kelp_timeseries_noNAs +
  geom_jitter() + # adds jitter to allow plotting of overlapping data points
  facet_wrap(~ SPECIES) # splits plots into individ. plots per species
```

# Part 6: Meta Questions

## Meta 1: 
### How did it feel working with an unfamiliar dataset? What did you notice?
    1. It took some time to understand what the variables were and what was actually collected in this data set
    2. For example, I had to do some research to understand what "percent cover" and "transect" mean
    2. A key (or variable guide) would be really helpful with any dataset I work with

## Meta 2:
### Split-Apply-Combine is…. a way of life, really. Is this something you have dealt with previously in your life or work? How comfortable are you with this concept?
    1. I have dealt with the concept a bit without knowing the name for it
    2. It's a bit of a logical pathway for problem solving
    3. I need to work on knowing when it is needed / when to apply it
    4. But I am comfortable with the use in this exercise

## Meta 3:
### When you’ve made datasets in the past, have they been wide, long, or something else? 
### After this week and the Browman and Woo paper, what advice would you give to future you when making data?
    1. Though I am new to the terms "wide" and "long", I have always made my datasets in the long format
    2. This is because, conceptually, it is easier for me to understand data when it is recorded by observation, with each column being a different variable / measurement
    3. When I have worked in Excel, I have ALWAYS made long datasets because filtering / sorting is really easy

## Meta 4:
### How much time did this take you, roughly? Again, I’m trying to keep track that these assignments aren’t killer, more than anything.
    1. This took me ~5 hours
    2. I had a bit of a hard time understanding what was being asked in some questions (ex: 4d, 5c)
    3. So I played around with these quite a bit to deduce the goal. In 4d, I just did double the work to address both ways I interpretted the question

## Meta 5:
### Please give yourself a weak/sufficient/strong assessment on this assigment. Feel free to comment on why.
    1. I will give myself a sufficient assessment
    2. I was able to figure out how to complete the exercises
    3. But, on 5c I chose to take a different solution that seemed easier. This might not have addressed the goal of the exercise but it saved a lot of time and effort
    2. I understand the concepts, but it took me quite a bit of effort and troubleshooting to apply them here
    3. I definitely need repeated practice to be able to do this without referencing the internet, my past work, class resources, etc.
