---
title: 'Homework 1: Intro to R!'
author: "Nathan Strozewski"
date: "2022-09-12"
output: html_document
---

## Part 1 ##  

### 1a: Load quakes with data(quakes)

```{r loading_quakes}  

data(quakes)  
quakes 

```

### 1b: Show what’s there with str() and summary(). What do you see?

```{r info_quakes}

str(quakes) 
summary(quakes) 

```

##### Notes:
    1. str() - variables are lat, long, depth, mag, stations
    2. summary() - min, max, median, quartiles for each variable

## Part 2 ##

### 2a: Show the entirety of the column long

```{r long_quakes}  

quakes[,"long"]

```

## Part 3 ##

### Part 3a: Hey, another useful function - unique()! 
    1. Apply it to a vector, and you can see what are all of the unique values ...
       It’s great for really digging into a problematic vector.<li>

```{r unique_quakes}

unique(quakes)
unique(quakes[,"long"])
unique(quakes[,"mag"])

```

### Notes:
    1. unique() returns the entered vector with duplicate data points remove
    2. unique() can be applied to an entire dataset or specific rows/cols
    3. I applied to cols here to test

## Part 4 ##

### Part 4a: What unique stations are there?
    1. Use length() with unique() to determine how many stations there are

```{r length_quakes}  
  
length(unique(quakes[,"stations"]))  
  
```

### Notes:
    1. I tested **?length** in the console to understand the new function
    2. I then established **(unique(quakes[,"stations"]))**
    3. I then applied **length(x)** where **x = (unique(quakes[,"stations"]))**
    4. Result: There are **102** unique stations in the quakes dataset

## Part 5 ##

### Part 5a: Using range(), what is the range of depths where quakes occur?

```{r range_quakes}

range(quakes[,"depth"])

```

### Notes:
    1. I tested **?range** in the console to understand the new function
    2. I then applied **range(x)** to **dataset[row,col] ...**
    3. where dataset is quakes, row is all (,) and col is "depth"
    4. Result: **40 - 680**

## Part 6 ##

### Part 6a: Where was the earthquake of largest magnitude found? 
    You’ll need to use some comparisons and the max() function for this ...
    in your indices instead of a number!

```{r which.max_quakes}

quakes[which.max(quakes[,"mag"]),]

```

### Notes:
    1. I first tested max on "mag" variable to know what value I was looking for ... 
    **max(quakes[,"mag"]) returned = 6.4**
    2. I then looked into **?which.max** to understand its usage
    3. Then I applied which.max to locate the index of the max "mag" value ...
    **which.max(quakes[,"mag"]) returned 152**
    4. I then pulled index 152 from the quakes dataset ...
    to see the lat, long, depth, and station values at the max ...
    **quakes[152,]** returned the: ...
    **lat, long, depth, and station values at mag 6.4**
    5. I cross-checked the mag (6.4) with max(quakes[,"mag"]) and it matched
    6. Lastly, I wanted to condense all these lines of code into one line
    7. I indexed quakes, with the row being the which.max line of code ...
    and the column being all (,)
    8. I cross-checked the return with that from quakes[152,] and it matched

## Part 7 - Meta Questions ##

### Meta 1.a.: How did it feel to connect concepts to a novel set of problems.
    1. This was incredibly helpful and hopefully how the course will continue
    2. Applying the concepts and skills helped me to better understand them
    3. I made many mistakes along the way that I learned from

###  Meta 1.b.: If you have experience with R, was this a cakewalk or challenging?
    1. I have dabbled with R for "fun" and have used MATLAB fairly extensively
    2. The biggest learning curve is switching over languages in my head
    3. I am not multi-lingual, but I imagine it is a similar mental switch
    4. For example, I keep typing % to comment instead of #
    5. Little mistakes like that add a lot of time and detract from focus
    6. I feel comfortable following the logic of R coding but do always need to check whether what I type is logical
    7. The biggest challenge is not knowing what options (i.e., functions) exist to solve the problem i am working on
    8. But I guess that's what Google and stack overflow are for!
    9. It's great that R detects error in code and suggests alternatives

### Meta 1.c.: How did it feel to shake off the cobwebs?
    1. Great! I am really enjoying this class
    2. It is great to take a class solely dedicated to developing a skill
    3. Rather than building a knowledge base
    4. I am very excited to continue
    5. The litany of resources + degree of organization for this class is very helpful

### Meta 2.: How much time did this take you, roughly?
    1, Approx. 1 hour for code, 30 min for markdown
    2. I really took my time to read into each function that I used
    3. Spend more time now to understand = spend less time on misunderstandings later!

### Meta 3.: Please give yourself a weak/sufficient/strong assessment
    1. I performed strongly on this assignment
    2. Good recall of the material we covered in class/lab
    3. Took time to read into each function
    4. Tested multiple ways to apply them
    5. Once I finished the question, I tried to consolidate the code into as few lines, functions, etc. as possible to improve my understanding, logic, and efficiency